Viscous Transport in Eroding Porous Media

### 2019-09-05 docs/transport/erosion_transport.tex
small changes to results

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points and a time step size of $\Delta t = 10^{-4}$.  The background
flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 2.  Snapshots of the bodies at four equispaced
times are shown in Figure~\ref{fig:Eroding20vort}.  The color is the
vorticity which is equivalent to the shear stress when restricted to the
eroding bodies.  Therefore, the rate of erosion is fastest in regions
where the magnitude of the vorticity is largest.  Initially, several of
the eroding bodies are closer to the outer wall than the $5h$ threshold
required to guarantee that the trapezoid rule achieves machine
precision.  In particular, with $h=L_2/N_\out$, the distance between
bodies 1, 6, 13, and 15 and the outer wall are $1.3h$, $2.9h$, $2.8h$,
and $1.3h$, respectively.  In addition, the distance between several
pairs of eroding bodies is too small to be accurately resolved with the
trapezoid rule at the prescribed resolution.  Particular pairs include
bodies 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18.  By using the Barycentric
quadrature rule, the interaction between these nearly-touching bodies
can be accurately resolved at the prescribed resolution, and erosion can
be simulated until all the bodies have vanished.

@@

In Figure~\ref{fig:Eroding20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  The spreading is
computed at several geometries that are formed during the erosion
process.  So that the spreading reaches a statistical equilibrium, we
use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories.  For all the reported
porosities, the particle dispersion exhibits two power low regimes.
Initially, the dispersion is ballistic ($\sigma_\lambda \sim t$) since
individual fluid particles have not yet explored enough space to
significantly alter their velocity.  However, once the particles have
been subjected to a range of velocities, their dispersion slows down,
and we observe super-diffusive ($\sigma_\lambda \sim t^\alpha$, $\alpha
\in (1/2,1)$) behavior over at least one order of magnitude in time.
Before any erosion takes place, the anomalous dispersion coefficient is
$\alpha = 0.56$.  Then, as the bodies begin to erode, the dispersion
rate grows towards the ballistic regime that corresponds to the absence
of bodies.  The monotonic increase in dispersion with respect to the
porosity is explained by the onset of channels where many tracers
experience less variability in their velocities.

@@

In Figure~\ref{fig:ErodingLow20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  As in the last
example, we analyze the spreading at several different
porosities and we use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories.  For all the reported
porosities, the particle dispersion exhibits two power low regimes, with
the lowest porosity geometry, $\phi=37.68\%$ having the clearest
transition.  Again, the dispersion transforms from a ballistic to
super-diffusive, and the anomalous disperion rate increases with the
porosity.  For this example, the dispersion is nearly Fickean at the
smallest porosity.
\todo[inline]{BQ: This paragraph has some false statements.  Need to
look closer at the data.  Might have accidently been looking at the
other 20 body example ie. Fickean dispersion}

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian20b}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/20b_dense_second_moment_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:ErodingLow20Transport} (a): The tortuosity of an
eroding geometry initialized with 20 bodies.  When compared to the last
example, the bodies are initially much closer together, and the porosity
is smaller.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue
dots).  The dash line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$
with $p=0.1669$, and the root mean square deviation is $1.13 \times
10^{-2}$.  (b): The temporal evolution of $\sigma_\lambda$ as a function
of time at eight different porosities.  The dashed line, which has slope
one, corresponds to the ballistic dispersion $\sigma_\lambda \sim t$ if
no bodies were present.  At early times, the particles undergo a
ballistic motion, but once they have traveresd a few grains, the
dispersion becomes super-dispersive with $\sigma_\lambda \sim
t^{\alpha}$, $\alpha \in (1/2,1)$.  The dashed-dotted lines are lines of
best fit with slope $0.92$ ($\phi=95.00\%$), $0.87$ ($\phi=85.17\%$),
$0.87$ ($\phi=75.10\%$), $0.91$ ($\phi=65.02\%$), $0.91$
($\phi=55.08\%$), $0.72$ ($\phi=45.03\%$), $0.69$ ($\phi=35.05\%$), and
$0.92$ ($\phi=30.67\%$).}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{100 eroding bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As a final example, we consider 100 eroding bodies with an initial
porosity near 55\%.  Snapshots of the configurations and vorticity are
in Figure~\ref{fig:Eroding100vort}.

\begin{figure}[H]
 \begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =0.9\linewidth]{./figs/100b_50}
\caption{100 bodies and porosity = 55.54\% at t= 0.005}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =1.1\linewidth]{./figs/100b_100}
\caption{99 bodies and porosity = 62.98\% at t= 0.01}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =0.9\linewidth]{./figs/100b_150}
\caption{94 bodies and porosity = 72.22\% at t= 0.015}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =1.1\linewidth]{./figs/100b_200}
\caption{82 bodies and porosity = 83.37\% at t= 0.02}
\end{subfigure}
\caption{\label{fig:Eroding100vort} Erosion of 100 bodies with the
  fixed-pressure-drop condition. The color is the vorticity of fluid. In
  this simulation, we use $N_\iin = 256$ discretization points on the
  body, $N_\out = 1024$ points on the outer wall, a time-step of
  $\Dt=10^{-4}$, and smoothing parameters of $\eps= 15/256$ and
  $\sigma=10/256$.}
\end{figure}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points and a time step size of $\Delta t = 10^{-4}$.  The background
flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 2.  Snapshots of the bodies at four equispaced
times are shown in Figure~\ref{fig:Eroding20vort}.  The color is the
vorticity which is equivalent to the shear stress when restricted to the
eroding bodies.  Therefore, the rate of erosion is fastest in regions
where the magnitude of the vorticity is largest.  Initially, several of
the eroding bodies are closer to the outer wall than the $5h$ threshold
required to guarantee that the trapezoid rule achieves machine
precision.  In particular, with $h=L_2/N_\out$, the distance between
bodies 1, 6, 13, and 15 and the outer wall are $1.3h$, $2.9h$, $2.8h$,
and $1.3h$, respectively.  In addition, the distance between several
pairs of eroding bodies is too small to be accurately resolved with the
trapezoid rule at the prescribed resolution.  Particular pairs include
bodies 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18.  By using the Barycentric
quadrature rule, the interaction between these nearly-touching bodies
can be accurately resolved at the prescribed resolution, and erosion can
be simulated until all the bodies have vanished.

@@

In Figure~\ref{fig:Eroding20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  The spreading is
computed at several geometries that are formed during the erosion
process.  So that the spreading reaches a statistical equilibrium, we
use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories.  For all the reported
porosities, the particle dispersion exhibits two power low regimes.
Initially, the dispersion is ballistic ($\sigma_\lambda \sim t$) since
individual fluid particles have not yet explored enough space to
significantly alter their velocity.  However, once the particles have
been subjected to a range of velocities, their dispersion slows down,
and we observe super-diffusive ($\sigma_\lambda \sim t^\alpha$, $\alpha
\in (1/2,1)$) behavior over at least one order of magnitude in time.
Before any erosion takes place, the anomalous dispersion coefficient is
$\alpha = 0.56$.  Then, as the bodies begin to erode, the dispersion
rate grows towards the ballistic regime that corresponds to the absence
of bodies.  The monotonic increase in dispersion with respect to the
porosity is explained by the onset of channels where many tracers
experience less variability in their velocities.  For this example, the
dispersion is nearly Fickean at the smallest porosity.

@@

In Figure~\ref{fig:ErodingLow20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  As in the last
example, we analyze the spreading at several different porosities and we
use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories to observe an asymptotic
dispersion rate.  For all the porosities, the dispersion is much closer
to ballistic when compared to the results in
Figure~\ref{fig:Eroding20Transport}.  However, there are still clear
transitions from ballistic dynamics at early times to asymptotic
super-diffusive dynamics.  In contrast to the lower porosity initial
condition (Section~\ref{sec:Eroding20}), at early times the erosion
results in a decrease in particle spreading. In particular, once erosion
has eroded the first 5\% of the bodies, the particle spreading
transitions from $\sigma_\lambda \sim t^{0.92}$ to $\sigma_\lambda \sim
t^{0.69}$.  To explain this behavior, note that the velocity is fairly
uniform, albeit small, at the initial condition (see
Figure~\ref{fig:ErodingLow20vort}).  Then as the geometry erodes, the
ratio of the fast and slow regions increases.  Since super-diffusive
dispersion results from a tracers spending times in a both slow and fast
regions, this results in an increase in the amount of super-dispersion.
\todo[inline]{Still need to work on this and explain how it contrasts
from the other 20 body example}
\todo[inline]{Need a reference for last sentence}

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian20b}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/20b_dense_second_moment_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:ErodingLow20Transport} (a): The tortuosity of an
eroding geometry initialized with 20 bodies.  When compared to the last
example, the bodies are initially much closer together, and the porosity
is smaller.  The tortuosity is calculated using the Eulerian
method~\eqref{eqn:tortuosity2} (blue dots).  The dash line is the
fitting line $\widehat{T}(\phi)=\phi^{-p}$ with $p=0.1669$, and the root
mean square deviation is $1.13 \times 10^{-2}$.  (b): The temporal
evolution of $\sigma_\lambda$ as a function of time at eight different
porosities.  The dashed line, which has slope one, corresponds to the
ballistic dispersion $\sigma_\lambda \sim t$ if no bodies were present.
At early times, the particles undergo a ballistic motion, but once they
have traversed a few grains, the dispersion becomes super-dispersive
with $\sigma_\lambda \sim t^{\alpha}$, $\alpha \in (1/2,1)$.  The
dashed-dotted lines are lines of best fit with slope $0.92$
($\phi=95.00\%$), $0.87$ ($\phi=85.17\%$), $0.87$ ($\phi=75.10\%$),
$0.91$ ($\phi=65.02\%$), $0.91$ ($\phi=55.08\%$), $0.72$
($\phi=45.03\%$), $0.69$ ($\phi=35.05\%$), and $0.92$ ($\phi=30.67\%$).}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{100 eroding bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As a final example, we consider 100 eroding bodies with an initial
porosity near 55\%.  Snapshots of the configurations and vorticity are
in Figure~\ref{fig:Eroding100vort}.  Again, by resolving near-contacts
with the Barycentric quadrature rule, a stable simulation is possible
with only $N_\iin = 256$ discretization points.  We use a time step size
of $\Delta t = 10^{-4}.$

\begin{figure}[H]
 \begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =0.9\linewidth]{./figs/100b_50}
\caption{}
%\caption{100 bodies and porosity = 55.54\% at t= 0.005}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =1.1\linewidth]{./figs/100b_100}
\caption{}
%\caption{99 bodies and porosity = 62.98\% at t= 0.01}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =0.9\linewidth]{./figs/100b_150}
\caption{}
%\caption{94 bodies and porosity = 72.22\% at t= 0.015}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =1.1\linewidth]{./figs/100b_200}
\caption{}
%\caption{82 bodies and porosity = 83.37\% at t= 0.02}
\end{subfigure}
\caption{\label{fig:Eroding100vort} Erosion of 100 bodies with the
  fixed-pressure-drop condition. The color is the vorticity of fluid. In
  this simulation, we use $N_\iin = 256$ discretization points on the
  body, $N_\out = 1024$ points on the outer wall, a time-step of
  $\Dt=10^{-4}$.  Because of the large number of bodies, erosion creates
  many channels connecting the inlet to the outlet.}
\end{figure}

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/100b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100Transport} (a): The tortuosity of an
eroding geometry initialized with 20 bodies.  When compared to the last
example, the bodies are initially much closer together, and the porosity
is smaller.  The tortuosity is calculated using the Eulerian
method~\eqref{eqn:tortuosity2} (blue dots).  The dash line is the
fitting line $\widehat{T}(\phi)=\phi^{-p}$ with $p=0.1669$, and the root
mean square deviation is $1.13 \times 10^{-2}$.  (b): The temporal
evolution of $\sigma_\lambda$ as a function of time at eight different
porosities.  The dashed line, which has slope one, corresponds to the
ballistic dispersion $\sigma_\lambda \sim t$ if no bodies were present.
At early times, the particles undergo a ballistic motion, but once they
have traversed a few grains, the dispersion becomes super-dispersive
with $\sigma_\lambda \sim t^{\alpha}$, $\alpha \in (1/2,1)$.  The
dashed-dotted lines are lines of best fit with slope $0.92$
($\phi=95.00\%$), $0.87$ ($\phi=85.17\%$), $0.87$ ($\phi=75.10\%$),
$0.91$ ($\phi=65.02\%$), $0.91$ ($\phi=55.08\%$), $0.72$
($\phi=45.03\%$), $0.69$ ($\phi=35.05\%$), and $0.92$ ($\phi=30.67\%$).}
\caption[inline]{Fix this caption}
\end{figure}
